---
title: "Feature Selection"
output:
    html_document:
        df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DESeq2)
library(matrixStats)
library(glmnet)
library(Boruta)
library(UpSetR)
library(c060)


#--- Load DESeq objects
if(exists("snakemake")) {
  cat("Running with snakemake\n")
  #TODO
  # load(snakemake@input[["deseq_rdata"]])
  # load(snakemake@input[["clin_rdata"]])
} else {
  cat("Not using snakemake\n")
  load("01-sample_data.Rdata")  
  load("02-clinical_data.Rdata")
  load("05-filter_counts.Rdata")
  load("06-unsupervised.Rdata")
  outdat <- '10-feature_selection.Rdata'
  out_pdf <- '10-feature_selection/upset.pdf'
}

```

### Store selected variables in this list

```{r}
selected_vars <- list()
```

### Retrotranscriptome expression data

```{r message=F, warning=F}
countDat <- filt.herv
cat(sprintf('%d total variables\n', nrow(countDat)))

stopifnot(all(colnames(countDat) == rownames(samples)))
samples$clust.herv <- clust.df$clust.retro.k4

dds <- DESeq2::DESeqDataSetFromMatrix(countDat, samples, ~1)
dds <- DESeq2::DESeq(dds, parallel=T)
tform <- DESeq2::varianceStabilizingTransformation(dds, blind=FALSE)
```

### LRT

```{r message=F, warning=F}
p.cutoff <- 1e-3
lfc.cutoff <- NULL

dds_lrt <- DESeq2::DESeqDataSetFromMatrix(countDat, samples, ~ clust.herv)
dds_lrt <- DESeq2::DESeq(dds_lrt, test="LRT", reduced = ~ 1, parallel=T)
res_lrt <- results(dds_lrt, alpha=p.cutoff)
sig_lrt <- subset(res_lrt, padj < p.cutoff)

selected_vars$lrt <- rownames(sig_lrt)
```

Selected `r nrow(sig_lrt)` variables using the LRT. Parameters: p.cutoff = `r p.cutoff`.


##### Transform data matrix

```{r}
mat <- assay(tform)
cat(sprintf('%d total variables\n', nrow(mat)))
```


### Randomized LASSO regression with stabiliy selection

```{r warning=F}
weakness <- 0.1
error <- 0.05
pi_thr <- 0.6

set.seed(12345)
res.stabpath <- c060::stabpath(samples$clust.herv, t(mat),
                      weakness=weakness, 
                      mc.cores=1, # prevents reseeding RNG on each core
                      family="multinomial",
                      type.multinomial="grouped"
                      )
res.stabsel <- c060::stabsel(res.stabpath, error=error, pi_thr=pi_thr, type="pfer")
selected_vars$lasso <- names(res.stabsel$stable)
```


Selected `r length(selected_vars$lasso)` variables using randomized LASSO. Paramters: weakness = `r weakness`; error = `r error`; pi_thr = `r pi_thr`.

```{r}
x <- plot(res.stabpath)
```


### Boruta

```{r}
set.seed(12345)
bor.orig <- Boruta(x=t(mat), y=samples$clust.herv, doTrace=0, ntree=1000, maxRuns=1000)
print(bor.orig)
bor.model <- TentativeRoughFix(bor.orig)
print(bor.model)

selected_vars$boruta <- names(bor.model$finalDecision)[bor.model$finalDecision == 'Confirmed']
```

Selected `r length(selected_vars$boruta)` variables using Boruta.

### Upset

```{r echo=F,warning=FALSE,message=FALSE,error=FALSE}
upset(fromList(selected_vars), order.by = "freq")
```

##### Get binary matrix

```{r}
binmat <- fromList(selected_vars)
rn <- do.call(c, selected_vars)
rn <- rn[!duplicated(rn)]
rownames(binmat) <- rn
rm(rn)
```



### SVM

##### Cross validation function

```{r}
svm.cross <- function(x, y, k=5, nrep=100) {
  lapply(1:nrep, function(rep) {
    folds <- caret::createFolds(y, k=k)
    svm.res <-lapply(folds, function(fld) {
      train <- x[-fld, ]
      test <- x[fld, ]
      train.lab <- y[-fld]
      test.lab <- y[fld]
      
      svm.model <- e1071::svm(y ~ . , data=data.frame(y=train.lab, train))
      pred <- predict(svm.model, data.frame(test))
      cm <- caret::confusionMatrix(table(pred, test.lab))
      list(svm.model=svm.model, pred=pred, cm=cm)
    })
    svm.res
  })
}
```


```{r}
set.seed(12345)
```


#### LASSO selected variables

```{r}
n <- binmat %>% 
  rownames_to_column("rowid") %>% 
  filter_at(vars(c('lasso')), ~.==1) %>%
  .$rowid
mat.sel <- mat[n,]

rres.lasso <- svm.cross(t(mat.sel), samples$clust.herv, k=5, nrep=100)
acc.lasso <- mean(sapply(unlist(rres.lasso, recursive = F), function(l) l[[3]]$overall['Accuracy']))
```


LASSO feature selection has accuracy = `r sprintf('%.3f', acc.lasso*100)` using a `r nrow(mat.sel)` HERV signature


#### Boruta selected variables

```{r}
n <- binmat %>% 
  rownames_to_column("rowid") %>% 
  filter_at(vars(c('boruta')), ~.==1) %>%
  .$rowid
mat.sel <- mat[n,]

rres.boruta <- svm.cross(t(mat.sel), samples$clust.herv, k=5, nrep=100)
acc.boruta <- mean(sapply(unlist(rres.boruta, recursive = F), function(l) l[[3]]$overall['Accuracy']))
```

Boruta feature selection has accuracy = `r sprintf('%.3f', acc.boruta*100)` using a `r nrow(mat.sel)` HERV signature

#### LRT selected variables

```{r}
n <- binmat %>% 
  rownames_to_column("rowid") %>% 
  filter_at(vars(c('lrt')), ~.==1) %>%
  .$rowid
mat.sel <- mat[n,]

rres.lrt <- svm.cross(t(mat.sel), samples$clust.herv, k=5, nrep=100)
acc.lrt <- mean(sapply(unlist(rres.lrt, recursive = F), function(l) l[[3]]$overall['Accuracy']))
```

LRT feature selection has accuracy = `r sprintf('%.3f', acc.lrt*100)` using a `r nrow(mat.sel)` HERV signature

```{r}
tabs.lasso <- Reduce('+', lapply(unlist(rres.lasso, recursive = F), function(l) l$cm$table))

cat('\nClassification\n')
print(tabs.lasso)
cat('\nClassification (proportion)\n')
print(tabs.lasso / sum(tabs.lasso))

prec.lasso <- lapply(unlist(rres.lasso, recursive = F), function(l) l$cm$byClass[,'Precision']) %>% bind_rows
rec.lasso <- lapply(unlist(rres.lasso, recursive = F), function(l) l$cm$byClass[,'Recall']) %>% bind_rows

cat('\nPrecision\n')
colMeans(prec.lasso)

cat('\nRecall\n')
colMeans(rec.lasso)
```


```{r}
cm2 <- as.table(matrix(c(tabs.lasso[1,1] + tabs.lasso[1,2] + tabs.lasso[2,1] + tabs.lasso[2,2],
                         tabs.lasso[1,3] + tabs.lasso[1,4] + tabs.lasso[2,3] + tabs.lasso[2,4],
                         tabs.lasso[3,1] + tabs.lasso[3,2] + tabs.lasso[4,1] + tabs.lasso[4,2],
                         tabs.lasso[3,3] + tabs.lasso[3,4] + tabs.lasso[4,3] + tabs.lasso[4,4]
                         ), nrow=2, ncol=2, byrow=T)
                )
colnames(cm2) <- c('C12','C34')
rownames(cm2) <- c('C12','C34')

caret::confusionMatrix(cm2)

```

```{r}
save(selected_vars,
     dds_lrt, res_lrt, sig_lrt,
     mat, samples,
     res.stabpath, res.stabsel,
     bor.orig, bor.model,
     rres.lasso, rres.boruta, rres.lrt,
     file=outdat
     )
```
